Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-07-08 main.tex
recovery intro Philippe
expl. algo best-search (en cours)

%\subsection{Best Search}
\label{sec:best}\label{sec:search}
\noindent
We shall now present a procedure for searching, for a~$\SWVPA$ $A$,
a word %$w \in \Omega^*$
of minimal weight for~$A$,
as stated in the following proposition. %\wrt~$\leq_\oplus$,
%
\begin{proposition}\label{th:best-search}
For a \SWVPA $A$
over $\Omega$,
$\Semiring$ commutative, bounded, total and complete, %semiring
and $\bar\Phi$ effective, %and $k$-convex, % label theory,
one can construct in PTIME a word $t \in \Omega^*$
such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$.
\end{proposition}
\florent{total?}
%
Let $A = \< Q, P, \init, \bar\wei, \final >$.
%
We propose a Dijkstra algorithm computing,
for every $q, q' \in Q$,
the minimum, \wrt $\leq_\oplus$, of the function
$\beta_{q, q'} : t \mapsto \weight_A(\config{q}{\bot}, t, \config{q'}{\bot})$.
Let us denote by $b_\bot(q, q')$ this minimum.
By definition of $\leq_\oplus$ it holds that:
%
\begin{equation}\label{eq:bbot}
  b_\bot(q, q') = \bigoplus_{t\in \Omega^*}
  \textstyle
  \weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr).
\end{equation}
Hence, following~\eqref{eq:SWVPA-value}, and the associativity and commutativity
and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is
\(
{\displaystyle\bigoplus_{t\in \Omega^*}}
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
\beta_{q, q'}(t)
\mathop{\otimes} \mathsf{out}(q')
=
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
b_\bot(q, q')
\mathop{\otimes} \mathsf{out}(q')
\).

\noindent
In order to compute

****

@@

\noindent
%More precisely,
We distinguish two cases : when the stack is empty, and when it is not.
\lydia{introduced 2 cases for b}
%
In the case of an empty stack, let $b_\bot : Q \times Q \to \Semiring$ be such that : %the function:
%
\begin{equation}\label{eq:bbot}
  b_\bot(q, q') = \bigoplus_{s\in \Omega^*}
  \textstyle
  \weight_A\bigl(\config{q}{\bot}, s, \config{q'}{\bot}\bigr).
\end{equation}
%
Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
and, providing that $\Semiring$ is total, it is the minimum in $\Omega^*$,
\wrt~$\leq_\oplus$, % this ordering.
of the fonction
$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$.
%
The term $\config{q}{\bot}, s, \config{q'}{\bot}$
of this sum is the central expression in
the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
of the function $\weight_A$.
\lydia{so ?}

%
\noindent
If the stack is not empty, let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
\florent{$b_\top$ : mot bien parenthèsé $c$/$r$}
and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
for every two states $q, q' \in Q$
and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$,
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*}
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
\end{equation}
%
Intuitively, the function defined in \eqref{eq:btop}
associates to $c \in \Omegac$
the minimum weight of a computation of $A$
starting in state $q$ with a stack
$\<c, p> \cdot \gamma \in \Gamma^+$
and ending in state $q'$ with the same stack,
such that the computation can not pop
the pair made of $c$ and $p$ at the top of this stack,
but may only read these symbols.
Moreover, $A$ may push another pair $\< c', p'>$ %call symbols
on the top of $\<c, p> \cdot \gamma$,
following the third case of
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol).
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
% Note that having a stack reduced to such a symbol makes impossible the application of the
% two last cases in the definition of $\weight_A$ (return symbol and empty stack).
% However, it is possible to apply the two first cases
% (internal symbol or call symbol, with a push on the top of $\top$).

>>>>

%\subsection{Best Search}
\label{sec:best}\label{sec:search}
\noindent
We shall now present a procedure for searching, for a~$\SWVPA$ $A$,
a word %$w \in \Omega^*$
of minimal weight for~$A$,
as stated in the following proposition. %\wrt~$\leq_\oplus$,
%
\begin{proposition}\label{th:best-search}
For a \SWVPA $A$
over $\Omega$,
$\Semiring$ commutative, bounded, total and complete, %semiring
and $\bar\Phi$ effective, %and $k$-convex, % label theory,
one can construct in PTIME a word $t \in \Omega^*$
such that $A(t)$ is minimal \wrt the natural ordering for $\Semiring$.
\end{proposition}
%\florent{total?}
%
Let $A = \< Q, P, \init, \bar\wei, \final >$.
%
We propose a Dijkstra algorithm computing,
for every $q, q' \in Q$,
the minimum, \wrt $\leq_\oplus$, of the function
$\beta_{q, q'} : t \mapsto \weight_A(\config{q}{\bot}, t, \config{q'}{\bot})$.
Let us denote by $b_\bot(q, q')$ this minimum.
By definition of $\leq_\oplus$, and since $\Semiring$ is total,
it holds that:
%
\begin{equation}\label{eq:bbot}
  b_\bot(q, q') = \bigoplus_{t\in \Omega^*}
  \textstyle
  \weight_A\bigl(\config{q}{\bot}, t, \config{q'}{\bot}\bigr).
\end{equation}
Since $\Semiring$ is complete, the infinite sum in \eqref{eq:bbot} is well defined,
and,  it is the minimum in $\Omega^*$,
\wrt~$\leq_\oplus$, % this ordering.
of the fonction
$s \mapsto \weight_A(\config{q}{\sigma}, s, \config{q'}{\sigma})$.
%
Hence, following~\eqref{eq:SWVPA-value}, and the associativity and commutativity
and distributivity for~$\otimes$ and~$\oplus$, the minimum of~$A(t)$ is
\(
{\displaystyle\bigoplus_{t\in \Omega^*}}
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
\beta_{q, q'}(t)
\mathop{\otimes} \mathsf{out}(q')
=
{\displaystyle\bigoplus_{q, q' \in Q}} \textstyle
\mathsf{in}(q) \mathop{\otimes}
b_\bot(q, q')
\mathop{\otimes} \mathsf{out}(q')
\).

\noindent
In order to compute the above function $b_\bot : Q \times Q \to \Semiring$, 
we shall use the auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
%
Intuitively, the function defined in \eqref{eq:btop}
associates to $c \in \Omegac$ 
the minimum weight of a computation of $A$
starting in state $q$ with a stack 
$\<c, p> \cdot \gamma \in \Gamma^+$ 
and ending in state $q'$ with the same stack,
such that the computation can not pop 
the pair made of $c$ and $p$ at the top of this stack,
but may only read these symbols.
Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
on the top of $\<c, p> \cdot \gamma$,
following the third case of 
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
% Note that having a stack reduced to such a symbol makes impossible the application of the 
% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
% However, it is possible to apply the two first cases 
% (internal symbol or call symbol, with a push on the top of $\top$).

@@

\noindent
%More precisely,
We distinguish two cases : when the stack is empty, and when it is not.
\lydia{introduced 2 cases for b}
%
In the case of an empty stack, let $b_\bot : Q \times Q \to \Semiring$ be such that : %the function:
%
EQBBOT
%
The term $\config{q}{\bot}, s, \config{q'}{\bot}$
of this sum is the central expression in
the definition \eqref{eq:weightA} of $A(s_0)$, for the minimum $s_0$
of the function $\weight_A$.
\lydia{so ?}

%
\noindent
If the stack is not empty, let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
\florent{$b_\top$ : mot bien parenthèsé $c$/$r$}
and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
for every two states $q, q' \in Q$
and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$,
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*}
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
\end{equation}
%
